---
title: "Project Report on Open Psychometric Data Analysis"
author: "Charles Yang, Bennet Brosseau, Carter Wunsch, Nicholas Heger, Isaac Tegeler"
date: "11/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

# Data was stored in CSV (Comma Delimited), but was tab delimited.
data <- read.table("VIQT_data.csv", sep="\t", header=TRUE)
originalData <- data
```

# Background and Questions Raised
**Psychological traits may point toward a stronger vocabulary. This has applications for teaching and further research**

1. How do answers to each supplemental question relate to the vocabulary?
2. How do demographics affect the vocabulary?
3. Which questions were the hardest?
4. How does childhood living region correlate with vocabulary?
5. Which supplemental questions were similar to others?
6. How does education level affect vocabulary?

# Data Acquisition/Preparation/Wrangling

* Data imported from [Open Psychometrics](http://openpsychometrics.org/_rawdata/VIQT_data.zip)
* .csv file was tab-delimited, so was read into R like a .tsv
* Question answers were in binary, so they were treated as their decimal equivalent
* Changed several columns to numeric type
* Demographics mapped from their code to actual value (IE: Urban, Education)
* Created and merged separate table with correct answers to each vocab question

# Visualization

## Heatmap for question 5
#### We can see some supplemental questions correlate with each other, meaning we can group them by personality indications
```{r}
x <- data %>% select(S1:S30)

c <- cor(x)
c <- data.frame(Var1 = row.names(c), c)
c <- c %>% pivot_longer(cols=S1:S30, names_to="Var2", values_to="value")

ggplot(c,aes(x=Var1, y=Var2, fill=value)) + geom_tile() + scale_fill_distiller(palette='YlGn')
```

## Bargraphs, Lineplots, Boxplot for questions 1, 2, 4, and 6
#### People who do the bare minimum to get by tend to have lower scores
```{r}
data %>%
  group_by(S26) %>%
  summarise(`Full Score`=mean(score_full), .groups="drop") %>%
  filter(S26 > 0) %>%
  ggplot(aes(x=S26, y=`Full Score`)) +
  geom_bar(stat="identity", width=.7) +
  ggtitle("Mean Full Scores by S26 Response")
```

#### It is clear that more educated individuals tend to have higher vocabulary.
```{r}
data$education <- recode(data$education, `1`="< HS", `2`="HS", `3`="UNI", `4`="GRAD") %>% factor(levels=c("< HS", "HS", "UNI", "GRAD"))
data %>%
  group_by(education) %>%
  summarise(`Full Score`=mean(score_full), .groups="drop") %>%
  ggplot(aes(x=education, y=`Full Score`)) +
  geom_bar(stat="identity", width=.7) +
  ggtitle("Mean Full Scores by Education")
```

#### Native English speakers clearly performed better in an English vocabular test.
```{r}
group_by(data, engnat) %>%
  summarise(`Full Score` = mean(score_full), .groups = 'drop') %>%
  filter(engnat > 0) %>%
  mutate(engnat = ifelse(engnat == 1, "Yes", "No")) %>%
  ggplot(aes(x = engnat, y = `Full Score`, fill = engnat)) + 
  geom_bar(stat = 'identity') +
  ggtitle("Mean Full Scores by Native English Speakers") +
  xlab("Native English Speaker")
```

#### There doesn't seem to be much of a relationship with Gender and Vocabulary
```{r}
dataGender <- data
dataGender$gender <- as.factor(dataGender$gender)
levels(dataGender$gender)[levels(dataGender$gender)==0] <- 'Did not answer'
levels(dataGender$gender)[levels(dataGender$gender)==1] <- 'Male'
levels(dataGender$gender)[levels(dataGender$gender)==2] <- 'Female'
levels(dataGender$gender)[levels(dataGender$gender)==3] <- 'Other'
ggplot(dataGender, aes(x=gender, y=score_full)) + geom_boxplot() + ggtitle('Gender vs Full Score') + xlab('Gender') + ylab('Full Score')
```

#### Age makes you wiser, at least in terms of vocabulary
```{r}
data %>%
  group_by(age) %>%
  summarise(`Full Score`=mean(score_full), .groups="drop") %>%
  filter(age < 100) %>%
  ggplot(aes(x=age, y=`Full Score`)) +
  geom_line(stat="identity", width=.7) +
  ggtitle("Mean Full Scores by Age")
```

## Linear regression for question 1
```{r}
data %>%
  group_by(S5) %>%
  summarise(`Full Score`=mean(score_full), .groups="drop") %>%
  filter(S5 > 0) %>%
  ggplot(aes(x=S5, y=`Full Score`)) +
  geom_point() +
  stat_smooth(method = "lm", col = "red")
```

## Tables for question 2
```{r}
answers <- read.table("answers.tsv", sep="\t", header=TRUE)
answers %>% pivot_wider(names_from = item, values_from = ans) -> answers

data %>% select(1:45, score_full) %>% arrange(-score_full) %>% top_frac(.5, score_full) -> topHalf
data %>% select(1:45, score_full) %>% arrange(-score_full) %>% top_frac(-.5, score_full) -> botHalf

discrimIdx <- data.frame(Question=character(), Discrimination=numeric())

#print("DISCRIMINATION INDICES BY QUESTION")
for (i in 1:45) {
  discrimIdx <- rbind(discrimIdx, data.frame(Question=paste(sep="","Q", i), Discrimination=as.numeric(table(as.numeric(topHalf[,i]) == as.numeric(answers[1,i]))[2]) / nrow(topHalf) - as.numeric(table(as.numeric(botHalf[,i]) == as.numeric(answers[1,i]))[2]) / nrow(botHalf)))
  
  #paste(sep="","Q", i, " ", as.numeric(table(as.numeric(topHalf[,i]) == as.numeric(answers[1,i]))[2]) / nrow(topHalf) - as.numeric(table(as.numeric(botHalf[,i]) == as.numeric(answers[1,i]))[2]) / nrow(botHalf)) %>% print
  
}
discrimIdx %>% arrange(-Discrimination)
```

# Answers to Questions Raised

1. Most did not correlate with vocab scores
  + S5, 21, 25, 26, 14, 16, 27, 10, 11, 20 had significant correlations
2. Demographics
  + Native english speakers performed better
  + Urban areas have greater variance in scores, with more in the lower ranges
  + Rural and Suburban have similar distributions
  + Age correlates postiively, and the trend is logarithmic, as it plateaus as people get older
  + Both male and Female had signfiicant outliers
    - Respondents who did not answer the gender question performed better
3. Questions 42, 21, and 41 were the hardest
4.Urban locations had more scores in a lower range than Rural. Suburban locations had double the values, but siilar distribution as Rural
5. Correlations between supplemental questions
  + Positive: S2 & S4, S6 & S12, etc.
  + Negative: S20 & S26, S18 & S26, etc.
6. Education correlates positively with scores

# Conclusion
### Correlation of Personalities and Vocabulary
* Low concientiousness correlates with lower final score, and vice versa
* Openness to experience correlates with a higher final score, and vice versa
* Neuroticism correlates with a lower final score

### Demographics
* Age, Education, English as a Native Language correlate positively with vocabulary
* Non-Urban individuals tend to perform better
* No significant differences between the two genders


**Raise kids to be concientious, open to new experiences, and in control of their emotinos to maximize their chance of having a strong vocabulary**