---
title: "HW4"
author: "Charles Yang"
date: "9/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(nycflights13)
```

# Are there any duplicate flight numbers?
```{r}
flights$flight <- factor(flights$flight)

group_by(flights, time_hour)$flight %>% duplicated() %>% any()
```

### Notes
* Some flight numbers are repeated.

# Add Origin and Destination to flights.
```{r}
airports %>% 
  select(origin_code=faa, origin_name=name, origin_lat=lat, origin_lon=lon) %>%
  right_join(y=flights, by = c("origin_code"="origin")) -> myFlights

airports %>% 
  select(dest_code=faa, dest_name=name, dest_lat=lat, dest_lon=lon) %>%
  right_join(y=flights, by = c("dest_code"="dest")) -> myFlights

head(myFlights)

```

# How does visibility affect delays?
## Departure Delay
```{r}
nycflights13::flights %>%
  left_join(y=weather %>% select(origin, visib, time_hour),
            by=c("time_hour", "origin")) -> weatherFlights

weatherFlights %>%
  group_by(visib) %>%
  summarize(dep_delay_median = median(dep_delay, na.rm = TRUE),
            dep_delay_mean = mean(dep_delay, na.rm = TRUE),
            .groups="keep")
```

## Arrival Delay
```{r}
weatherFlights %>%
  group_by(visib) %>%
  summarize(arr_delay_median = median(arr_delay, na.rm = TRUE),
            arr_delay_mean = mean(arr_delay, na.rm = TRUE),
            .groups="keep")
```

### Notes
* Higher visibility tends to reduce delays, and vice versa.

# Computing and Creating Delay Severity Column
```{r}
myFlights %>%
  mutate(severe=if_else(!is.na(arr_delay), if_else(arr_delay >= 30, 1, 0), 0)) %>%
  group_by(dest_code) %>%
  summarize(delay_severity = as.double(sum(severe)) / n(), .groups="keep") -> delaySeverity_by_faa

delaySeverity_by_faa %>% arrange(-delay_severity) %>% head(3)
```

### Notes
* Columbia, Tulsa, and Oklahoma City had the highest Delay Severity

# Delay Severity by Airport on June 13, 2013
```{r}
myFlights %>%
  filter(month==6, day==13) %>%
  mutate(severe=if_else(!is.na(arr_delay), if_else(arr_delay >= 30, 1, 0), 0)) %>%
  group_by(dest_code) %>%
  summarize(delay_severity = as.double(sum(severe)) / n(), .groups="keep") %>%
  arrange(-delay_severity)
```

### Notes
* This was the time of the 2-13 Derecho Series, where two massive storm systems impacted almost the entirety of Eastern United States.
* The extreme delays were caused by this storm.

