---
title: STAT 477 HW6
author: Charles Yang
date: 3/20/2022
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

```{r setup, include=FALSE}
# knitr::opts_chunk$set(results = "hide")
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

library(tidyverse)

library(ggmosaic)
library(vcdExtra)

correlation_coefficient <- function(table) {
  chisqstat <- chisq.test(table, correct = F)$statistic
  n <- sum(table)
  sgn <- sign(table[1, 1] * table[2, 2] - table[1, 2] * table[2, 1])
  sgn * sqrt(chisqstat / n)
}

crammer_v <- function(table) {
  chisqstat <- chisq.test(table,
    correct = F
  )$statistic
  min.dim <- min(dim(table))
  numobs <- sum(table)
  sqrt((chisqstat / numobs) / (min.dim - 1))
}


tires <- read.csv("tires.csv")
names(tires)[2] <- "perfect"

norc <- read.csv("norc.csv")
norc$Income <- factor(norc$Income, levels = c("Below", "Average", "Above"))
norc$Happy <- factor(norc$Happy, levels = c("Not Too", "Pretty", "Very"))
```

# Problem 1: Gun Registration & Death Penalty
### Test of Independence
```{r}
opinions <- matrix(c(784, 236, 311, 66),
  nrow = 2,
  ncol = 2,
  byrow = TRUE,
)

opinions_independence <- chisq.test(opinions)
opinions_independence
```

I tabulated the opinions in R and ran a chi-squared test.
It calculates the joint probability of all pairs of categories and the product of their respective marginal probabilities.
Then it tests if those are equal. If the data suggests they aren"t equal, then we get a lower p-value.
There is strong envidence that the opinions are not independent.

### Correlation Coefficient
```{r}
opinions_correlation <- correlation_coefficient(opinions) %>% as.double
opinions_correlation
```

A tiny negative correlation. Seems uncorrelated.

----------

# Problem 2: Tires
### Test of Independence with Grouping Variable
Whether or not a variable is a grouping variable does not matter.
We can still test for independence because it is categorical.
Addtionall, we may want to see if tire condition is dependent on the shift.

### Alternative Universe and Data/Test
Suppose we wanted to test if there was a difference in the proportion of tires that were perfect between the shifts.
In this case, we would test for equality of these proportions.

### Contigency Table & Mosaic Plot
```{r}
ggplot(data = tires) +
  geom_mosaic(aes(x = product(perfect, shift), fill = perfect)) +
  labs(
    x = "Shift",
    y = "Perfect",
    title = "Tires Perfect by Shift"
  )

tires_contigency <- table(tires$shift, tires$perfect)
tires_contigency
```

The number of tires decreases as shift increases.
Shifts 1 & 2 seem to have 50/50 perfect tires, but shift 3 has less perfect tires.

### Test of Independence
```{r}
tires_independence <- chisq.test(tires_contigency)
tires_independence
```

H0: The tires" conditions are independent of the shift. <br>
H1: The tires" conditions are not independent on the shift. <br>

Chi-squared test statistic: 5.2942, df = 2 <br>
p-value: 0.07085 <br>

There is moderate strength evidence that the tires" conditions are not independent of the shift.

### Correlation Coefficient & Crammer V
```{r}
tires_correlation <- correlation_coefficient(tires_contigency) %>% as.double
tires_correlation

tires_crammer <- crammer_v(tires_contigency) %>% as.double
tires_crammer
```

----------

# Problem 3: Perceived Happiness & Income
### Contigency Table
```{r}
norc_contigency <- table(norc$Income, norc$Happy)
norc_contigency
```

### Test of Independence
```{r}
norc_independence <- chisq.test(norc_contigency)
norc_independence
```

### Concordant with Very Happy & Average Income
1. Pretty Happy - Below Income
2. Not Too Happy - Below Income

### Disconcordant with Pretty Happy - Below Average Income
1. Not Too Happy - Above Income
2. Not Too Happy - Average Income

### Goodman-Kruskal Gamma
```{r}
GKgamma(norc_contigency)
```

There appears to be a weak positive-directional relationship between happiness and income level.