---
title: STAT 477 HW1
author: Charles Yang
date: 1/18/2022
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

```{r setup, include=FALSE}
# knitr::opts_chunk$set(results = "hide")
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(plyr)
library(ggplot2)
```


# 1. Examples of Categorical RVs
1. Education
    + Ordinal, ordered by level
    + Example: GED, High School, Bachelor's Degree, Master's Degree, Doctorate
2. Approach Difficulty (in rock climbing)
    + Ordinal, ordered by difficulty of approach
    + Example: Class 1, Class 2, Class 3, Class 4, Class 5
3. Rock Climbing Discipline
    + Nominal
    + Example: Lead, Speed, Bouldering, Ice, Mixed
4. Climbing Style
    + Nominal
    + Example: Static, Dynamic, Unorthodox
5. Climbing Hold Type
    + Nominal
    + Example: Jug, Crimp, Sloper, Pinch, Pocket, Volume, Aid

<hr>

# 2. STAT 477 Spring 2022 Data

```{r load_data}
#setwd("stat477/hw1")
class_spring_22 <- read.csv("SPRING2022STAT_477_577.csv", header = T) # nolint
class_spring_22 <- dplyr::rename(class_spring_22, Classification = Clsfn.Year)
```

```{r}
classes_summary <- count(class_spring_22$Classification)
classes_summary$rel <- round(prop.table(classes_summary[2]), 4)
classes_summary
```

```{r}
ggplot(data = class_spring_22, aes(x = Classification)) +
  geom_bar(fill = "purple") +
  labs(x = "Classification", y = "Number of Students") +
    ggtitle("STAT 477 Spring 2022 Students by Classification")
```

```{r}
pie(t(classes_summary[2]), labels = t(classes_summary[1]),
  col = c("blue", "aquamarine", "green"),
  main = "STAT 477 Spring 2022 Students by Classification")
```

## Year Classification
* Most students are year 4 classification
* There are a significant number of year 3s
* There are a few graduate students
* Besides missing absolute values, the pie chart is nice here

```{r}
majors_summary <- count(class_spring_22$Major)
majors_summary$rel <- round(prop.table(majors_summary[2]), 4)
majors_summary
```

```{r}
ggplot(data = class_spring_22, aes(x = Major)) +
  geom_bar(fill = "gold") +
  labs(x = "Major", y = "Number of Students") +
    ggtitle("STAT 477 Spring 2022 Students by Major")
```

```{r}
pie(t(majors_summary[2]), labels = t(majors_summary[1]),
  main = "STAT 477 Spring 2022 Students by Major")
```

## Major
* Nearly half of the students are DS
* The number of STAT majors is about half that of DS
* There is an assortment of other majors with 1-2 students
* The pie chart is hard to interpret because there are so many majors.
* Additionally, it is hard to tell absolute differences between categories.