---
title: STAT 477 HW5
author: Charles Yang
date: 3/6/2022
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(ggmosaic)

titanic <- read.csv("Titanic.csv")
titanic$Class <- factor(titanic$Class, levels = c("First", "Second", "Third", "Crew"))
titanic$Status <- factor(titanic$Status)

GSS <- read.csv("GSS.csv")
GSS$Wrong <- factor(GSS$Wrong, levels = c("Never", "Sometimes", "Almost.Always", "Always"))

aspirin <- read.csv("aspirin.csv")

# functions
rr_ci <- function(y, n, conf_level) {
  y1 <- y[1]
  y2 <- y[2]
  n1 <- n[1]
  n2 <- n[2]
  alpha <- 1 - conf_level
  z <- qnorm(1 - alpha / 2)

  phat1 <- y1 / n1
  phat2 <- y2 / n2

  rr <- phat1 / phat2

  selogrr <- sqrt((1 - phat1) / (n1 * phat1) + (1 - phat2) / (n2 * phat2))

  logrr.lower <- log(rr) - z * selogrr
  logrr.upper <- log(rr) + z * selogrr

  rr.lower <- exp(logrr.lower)
  rr.upper <- exp(logrr.upper)

  cat("Estimated Relative Risk = ", rr, "\n")
  cat("Confidence Interval for Population Relative Risk = ", rr.lower, rr.upper, "\n")
}

or_ci <- function(y, n, conf_level) {
  y1 <- y[1]
  y2 <- y[2]
  n1 <- n[1]
  n2 <- n[2]
  alpha <- 1 - conf_level
  z <- qnorm(1 - alpha / 2)

  phat1 <- y1 / n1
  phat2 <- y2 / n2

  or <- (phat1 / (1 - phat1)) / (phat2 / (1 - phat2))

  selogor <- sqrt(1 / (n1 * phat1) + 1 / (n1 * (1 - phat1)) + 1 / (n2 * phat2) + 1 / (n2 * (1 - phat2)))

  logor.lower <- log(or) - z * selogor
  logor.upper <- log(or) + z * selogor

  or.lower <- exp(logor.lower)
  or.upper <- exp(logor.upper)

  cat("Estimated Odds Ratio = ", or, "\n")
  cat("Confidence Interval for Population Odds Ratio = ", or.lower, or.upper, "\n")
}
```


# Problem 1: Titanic Outcome by Class
### Contingency Table & Mosaic Plot
```{r}
twoway <- table(titanic$Class, titanic$Status)
twoway

ggplot(data = titanic) +
  geom_mosaic(aes(x = product(Class, Status), fill = Class)) +
  labs(
    x = "Status",
    y = "Class",
    title = "Titanic Class by Status"
  )
```

The mosaic plot shows a much larger proportion of first and second class passengers who survived.

### Overall Hypothesis Test
```{r}
gn <- margin.table(twoway, 1)
prop.test(twoway[, 1], gn, alternative = "two.sided", correct = F)
```

H0: p1 = p2 = p3 = p4, HA: !(p1 = p2 = p3 = p4), X-squared = 187.79, p-value < 2.2e-16 <br>
There is overwhelming evidence that the proportion of passengers rescued is not the same across all classes.

### Pairwise Hypothesis Tests
```{r}
prop.table(twoway)

pairwise.prop.test(twoway[, 1], gn, p.adjust.method = "BH")
connected_letters <- matrix(c(
  "Third", sum(prop.table(twoway)[2, ]), "A", "", "",
  "Crew", sum(prop.table(twoway)[1, ]), "A", "", "",
  "Second", sum(prop.table(twoway)[3, ]), "", "B", "",
  "First", sum(prop.table(twoway)[4, ]), "", "", "C"
),
ncol = 5, byrow = TRUE
)
colnames(connected_letters) <- c("Class", "Sample Proportion", "Levels", "       ", "       ")
as.table(connected_letters)
```

It appears that third and crew had similar rescue rates.
Second class is different from all others, and so is First.
The perception from the movie was correct.

<hr>

# Problem 2: Premarital Sex Opinion by Religion
### Contingency Table and Mosaic Plot
```{r}
twoway <- table(GSS$Wrong, GSS$Religion)
twowayprop <- prop.table(twoway)
twoway

ggplot(data = GSS) +
  geom_mosaic(aes(x = product(Wrong, Religion), fill = Religion)) +
  labs(
    x = "Religion",
    y = "Wrong",
    title = "When is premarital sex wrong?"
  )
```

Protestants seem to be on the extremes only, where Always and Never are the most common.
The non-religious and jewish people have majorities that don't mind premarital sex is a problem.
Catholic and other religious groups are less chill, but disproval of premarital sex for all groups but protestants has a negative relationship with the number of people.  

### Conditional Distribution of Catholics
```{r}
twoway[, 1] / sum(twoway[, 1])
```

### Conditional Distribution of Protestants
```{r}
twoway[, 5] / sum(twoway[, 5])
```

### Overall Hypothesis Test
```{r}
gn <- margin.table(twoway, 1)
prop.test(twoway[, 1], gn, alternative = "two.sided", correct = F)
```

There is overwhelming evidence that the opinions on premarital sex between at least two religious groups are different.

### Fix expected value problem
I would combine collapse jewish with other religious groups since they are relatively low in numbers.

<hr>

# Problem 3: Aspirin and Heart Attacks
### Contingency Table and Mosaic Plot
```{r}
twoway <- table(aspirin$Heart.Attack, aspirin$Group)
twowayprop <- prop.table(twoway)
twoway

ggplot(data = aspirin) +
  geom_mosaic(aes(x = product(Heart.Attack, Group), fill = Group)) +
  labs(
    x = "Treatment",
    y = "Heart Attack",
    title = "Aspirin x Heart Attack"
  )
```

It appears that aspirin is more effective than placebo in reducing heart attack risk.

### Conditional Probabilities
```{r}
paste0("Aspirin: ", toString(twoway[, 1] / sum(twoway[, 1])))
paste0("Placebo: ", toString(twoway[, 2] / sum(twoway[, 2])))
```

### Overall Hypothesis Test
```{r}
gn <- margin.table(twoway, 1)
prop.test(twoway[, 1], gn, alternative = "greater", correct = F)
```

There is overwhelming evidence that aspirin is more effective than placebo in reducing heart attack risk.b 

### CI for Difference in Heart Attack Risk Between Groups
```{r}
diff <- 0.009 - 0.017
p <- length(aspirin$Heart.Attack[aspirin$Heart.Attack == "Yes"]) / length(aspirin$Heart.Attack)
z <- 1.96
se <- sqrt(p * (1 - p) / sum(twoway[, 1]) + p * (1 - p) / sum(twoway[, 2]))
ci <- c(diff - se * z, diff + se * z)
ci
```

We are 95% confident that the true difference in proportions of hearta attack between placebo and aspirin treatments is between -0.011 and -0.0050.
Apsirin is more effective than placebo in preventing heart attacks.

### Relative Risk
```{r}
rr_ci(twoway[, 1], gn, conf_level = 0.95)
```

We are 95% confident that the relative risk of heart attack when taking placebo instead of aspirin is between 1.21 and 1.65 times higher.

### Odds Ratio
```{r}
or_ci(twoway[, 1], gn, conf_level = 0.95)
```

We are 95% confident that the odds of heart attack when taking placebo instead of aspirin is between 1.44 and 2.33 times higher.