---
title: "ds202_hw2"
author: "Charles Yang"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```
# **R Commands**
## Convert F to C
```{r}
cel <- c((85 - 32) * 5 / 9, (67 - 32) * 5 / 9)
cel
```

## 200 -> 1, every third entry is 0
```{r}
vec200 <- 200:1
vec200[seq(3,200, by=3)] <- 0
vec200
```

## Perfect Squares from 1 -> 2001
```{r}
numPerfect <- 1:2001
numPerfect <- numPerfect[sqrt(numPerfect) == round(sqrt(numPerfect), digits = 0)]
numPerfect
```

# **Cars and MPG**
```{r}
names(mpg)
```

## 5 Cars with the worst highway MPG
```{r}
worst5 <- head(mpg[order(mpg$hwy),], 5)
worst5
```
### Notes:
* Four cars with the worst Mileage are Ford brand. The fifth is Jeep.
* They all have a city and highway mpg of 9 and 12, respectively.

## Number of unique pickup models
```{r}
numPickups <- length(table(subset(mpg, class=="pickup")$model))
numPickups
```
### Notes:
* There are 4 unqiue pickup models.

## Plot city vs highway MPG
```{r}
plot(mpg$hwy, mpg$cty)
```
### Notes:
* There is a strong positive correlation between highway mpg and city mpg.
* There are a limited amount of points because there are points with the same value hidden within eachother. Giving the points an opacity value would help visualize where this is occuring.

## MPG comparison between Ford and Toyota
```{r}
ford <- subset(mpg, manufacturer == "ford")
toyota <- subset(mpg, manufacturer == "toyota")
boxplot(ford$cty, ford$hwy, toyota$cty, toyota$hwy, main="Ford vs Toyota MPG",ylab="MPG", xlab="Manufacturer and MPG Type", names=c("Ford City", "Ford Highway", "Toyota City", "Toyota Highway"), col=c("red", "red", "grey", "grey"))
abline(h=mean(ford$cty), col="red")
abline(h=mean(ford$hwy), col="red", lty=2)
abline(h=mean(toyota$cty), col="grey")
abline(h=mean(toyota$hwy), col="grey", lty=2)
```
### Notes:
* From the graph, we can see Toyota has the highest possible MPG
* Ford's median MPG in city or highway is less than Toyota's
* Additionally, when comparing means, as denoted by the horizontal lines, we can see that Toyota outperforms Ford in city and highway MPG
* Overall, Toyota outperforms ford in MPG
