---
title: STAT 477 HW7
author: Charles Yang
date: 3/30/2022
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(irr)

mcnemar_ci <- function(table, conf_leve = 0.95) {
    alpha <- 1 - conf_leve
    z <- qnorm(1 - alpha / 2)

    n <- sum(table)
    y1_ <- margin.table(table, 1)[1]
    y_1 <- margin.table(table, 2)[1]
    diff12 <- table[1, 2] - table[2, 1]
    add12 <- table[1, 2] + table[2, 1]
    se <- sqrt(add12 - diff12^2 / n) / n
    lowerci <- y1_ / n - y_1 / n - z * se
    upperci <- y1_ / n - y_1 / n + z * se
    cat("Confidence Interval = ", lowerci, upperci, "\n")
}
```

# Problem 1: Baby Pictures
### McNemar's Test
```{r}
judgements <- matrix(
    c(
        12 + 43 + 14, 45 + 22 + 20,
        54, 36
    ),
    nrow = 2,
    ncol = 2,
    byrow = TRUE,
)

mcnemar.test(judgements, correct = F)
```

There is overhwelming evidence that the judges perform differently in identifying Dr. Floelich's baby when looking at a baby or adult picture of Dr. Foelich.

### McNemar's Confidence Interval
```{r}
mcnemar_ci(judgements)
```

We are 95% confident that the proportion of judges that can identify Dr. Foelich's baby is 4.1% to 22.7% more when looking at her adult picture compared to her baby picture.